---
title: "Lab 06 - Sad plots"
author: "Lindley Slipetz"
date: 4/16/2021
output: github_document
---

### Load packages and data

```{r load-packages, message = FALSE, warning = FALSE}
library(tidyverse) 
```

```{r load-data, message = FALSE, warning = FALSE}
staff <- read_csv("data/instructional-staff.csv")
```

```{r pivot_data}
staff_long <- staff %>%
  pivot_longer(cols = -faculty_type, names_to = "year") %>%
  mutate(value = as.numeric(value))
```

```{r line_plot_bad}
staff_long %>%
  ggplot(aes(x = year, y = value, color = faculty_type)) +
  geom_line()
```



### Exercise 1
```{r good_line_graph}
staff_long %>%
  ggplot(aes(x = year, y = value, group = faculty_type, color = faculty_type)) +
  geom_line() +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(x = "Year", y = "Percentage of Hires", title = "Percentage of hires by year grouped by faculty type", color = "Faculty type")
```

### Exercise 2

If I'm understanding this correctly, it should be a comparison of TT faculty grouped together compared to Part-Time faculty. I don't think combining them as is would make sense because you could potentially end up with values over 100% and it'd be kind of meaningless. Maybe take the average percent hires of TT faculty per year and the part time faculty, then do a line graph of the grouped average and the part-time faculty (I didn't include grad students because that didn't seem to be the question. I don't think they should be grouped with TT faculty because that's totally different, but they shouldn't be grouped with part-time because they're not faculty. A separate line could be included for grad students. I also didn't include fulltime non-TT because they're important different from TT). Okay, I think I'm going to do that.

```{r group_TT}
TT_staff <- staff_long %>%
  filter(faculty_type == "Full-Time Tenured Faculty" |
           faculty_type == "Full-Time Tenure-Track Faculty") %>%
  group_by(year) %>%
  mutate(avg_TT = mean(value)) %>%
  filter(faculty_type == "Full-Time Tenured Faculty")
PT_staff <- staff_long %>%
  filter(faculty_type == "Part-Time Faculty") %>%
  group_by(year) %>%
  mutate(avg_TT = value)
all_staff <- rbind(TT_staff, PT_staff)
```

Now we're going to do the graph.
```{r graph_2}
all_staff %>%
  ggplot(aes(x = year, y = avg_TT, group = faculty_type, color = faculty_type)) +
  geom_line() +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(x = "Year", y = "Percentage of Hires", title = "Percentage of hires by year grouped by faculty type") +
  scale_color_discrete(name = "Faculty Type", labels = c("TT", "Part-time"))
```

We see that TT faculty have steadily declined since 1975, while part time faculty have steadily increased. Now, nearly half of new hires are part time.


### Exercise 3

```{r fisheries_data}
fisheries <- read_csv("data/fisheries.csv")
```

Given the wide range of data, I think this would be best split into two graphs: the top 10% and bottom 10%. We'll start with the top 20.

```{r fish_desc}
fish_desc <- fisheries %>%
  arrange(desc(total))
```
```{r top_10}
obs <- nrow(fish_desc) 
top_10 <- fish_desc %>% 
  filter(row_number() < obs * 0.1)
```

```{r top_10_graph}
top_10 %>%
ggplot(aes(x=reorder(country, -total), y=total)) +
  geom_bar(stat="identity") +
  coord_flip() + 
  labs(title = "Fishery production of countries-2016", x = "Country", y = "Fishery production")
```

China is byfar the biggest fishery producer. It is so much larger than the other countries that it's no wonder the original graph was difficult to read. Another approach to tackling this data is the remove the outliers like China and Indonesia. We'll keep on track and look at the bottom 10%. 

```{r bottom_10}
bot_10 <- fish_desc %>%
    top_n(total
          ,n = -0.1*nrow(.))
```

```{r bot_10_graph}
bot_10 %>%
ggplot(aes(x=reorder(country, -total), y=total)) +
  geom_bar(stat="identity") +
  coord_flip() + 
  labs(title = "Fishery production of countries-2016", x = "Country", y = "Fishery production")
```

There are 4 countries with zero fish production. These could also be eliminated if we took the outlier approach. Lesotho has the most of the bottom 10%, but it is still much less than the amounts on the top 10% graph. Hence, I think this is best split into multiple graphs.